PHAT RAID, yo

### 2014-05-14 paper/paper.tex
spelling fix
line was too long
Add a theoretical analysis of failure probabilities

\begin{figure}
  \centering
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {3/1, 3/2, 3/3, 1/3, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{Progress, No Data Loss}
    \label{fig:failures_allok}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 2/1, 2/2}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, No Data Loss}
    \label{fig:failures_noprogress}
  \end{subfigure}
\\\vspace{1em}
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[scale=.5,shorten <=2pt, shorten >=2pt,]
      \node[rotate=25] (N) at (1.5,-1.5) {\textsc{Impossible!}};
    \end{tikzpicture}
    \caption{Progress, Data Loss}
    \label{fig:failures_max}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, Data Loss}
    \label{fig:failures_noprogress_nodata}
  \end{subfigure}

@@

In general, progress ceases if more than one cluster cannot reach consensus and
data loss occurs if more than one cluster completely fails. The lower bound of
failed nodes for progress cessation in \phatraidcf{$C$}{$f$} is:
    $$ C\cdot(f+1) $$
and the lower bound of failed nodes for data loss is:
    $$ 2\cdot(2f+1) $$

In contrats, if all data has been propagated to all nodes, data is
only lost when all nodes die. If data has not been propagated to all nodes, then
loss occurs if all nodes containing the data die.

@@

>>>>

\begin{figure}
  \centering
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {3/1, 3/2, 3/3, 1/3, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{Progress, No Data Loss}
    \label{fig:failures_allok}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 2/1, 2/2}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, No Data Loss}
    \label{fig:failures_noprogress}
  \end{subfigure}
\\\vspace{1em}
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[scale=.5,shorten <=2pt, shorten >=2pt,]
      \node[rotate=25] (N) at (1.5,-1.5) {\textsc{Impossible!}};
    \end{tikzpicture}
    \caption{Progress, Data Loss}
    \label{fig:failures_impossible}
  \end{subfigure}
~
  \begin{subfigure}{0.4\textwidth}
    \centering
    \begin{tikzpicture}[<->,scale=.5,shorten <=2pt, shorten >=2pt,]
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3, 3/1, 3/2, 3/3}
        \node [circle, draw, thick] (pr\x\y) at (\x,-\y) {};
      \foreach \x/\y in {1/1, 1/2, 1/3, 2/1, 2/2, 2/3}
        \node [strike out, draw=red, thick] (pr\x\y) at (\x,-\y) {};
    \end{tikzpicture}
    \caption{No Progress, Data Loss}
    \label{fig:failures_noprogress_dataloss}
  \end{subfigure}

@@

In general, progress ceases if more than one cluster cannot reach consensus and
data loss occurs if more than one cluster completely fails. The lower bound of
failed nodes for progress cessation in \phatraidcf{$C$}{$f$} is:
    $$ 2\cdot(f+1) $$
and the lower bound of failed nodes for data loss is:
    $$ 2\cdot(2f+1) $$

In contrast, if all data has been propagated to all nodes, data is only lost
when all nodes die. If data has not been propagated to all nodes, then loss
occurs if all nodes containing the data die.

\subsubsection{Theoretical Analysis}

A single number cannot describe the failure threshold of a \phatraid{} cluster
because of the interplay between VR and RAID. The failure of the \phatraid{}
group is dependent on the arrangement of the failed nodes. For example, Figure
\ref{fig:failures_noprogress} depicts a situation wherein only four failures
stops progress. In contrast, Figure \ref{fig:failures_allok} depicts a situation
wherein five failures does \emph{not} stop progress.

In a \phatraidcf{3,1} group, less than one-quarter of four-failure arrangements
stop progress. The necessary condition for progress cessation is that two or
more clusters have lost at least $f+1$ nodes. By the pidgeon hole principle,
only the two-in-two, four-failure arrangement stops progress. We can count the
number of such arrangements:

    $$ \binom{3}{2} \binom{3}{2}^2 = 3^3 = 27 $$

The first term represents choosing the two clusters of three that will
fail. The second term represents choosing which two nodes in each cluster of
three nodes will fail.

The total number of four-failure arrangements is simply:

    $$ \binom{9}{4} = 126 $$

If all nodes fail uniformly and independently, then the probability of four
nodes causing \phatraid{} group failure is

    $$ \frac{27}{126} \approx 0.21 $$

When five nodes fail, they can be distributed into each group in the
proportions: (3,1,1), (3,2,0) and (2,2,1). The first case does not stop progress
because only one cluster fails. The later two cases stop progress because two
clusters fail. We calculate the probability of \phatraid{} group failure when
five nodes fail:

\begin{align*}
  \textrm{(3,1,1)}\;\;\;
    &\binom{3}{1}\binom{3}{3}\binom{3}{1}\binom{3}{1} = 3^3 = 27 \\
  \textrm{(3,2,0)}\;\;\;
    &\binom{3}{1}\binom{2}{1}\binom{3}{3}\binom{3}{2} = 2\cdot3^2 = 18 \\
  \textrm{(2,2,1)}\;\;\;
    &\binom{3}{1}\binom{3}{2}\binom{3}{1}\binom{3}{1} = 3^4 = 81 \\
  \textrm{5-Failure Arrs.}\;\;\;
    &\binom{9}{5} = 126 \\
  \textrm{Failure Prob.}\;\;\;
    &\Pr\left[\textrm{group failure} \,|\, n_{failed} = 5\right] = \frac{81 + 18}{126} =
    \frac{99}{126} \approx 0.79 \\
\end{align*}

A VR or PAXOS cluster of size nine is resilient to $f = 4$ failures. In
contrast, \phatraidcf{3}{1} is absolutely resilient to $f=3$ failures, has high
probability of resilience to $f=4$ failures, and has low probability of
resilience to $f=5$ failures.

The general formula for the failure probability distribution is not clear, but
it is clear that a \phatraidcf{$C$}{$f$} group is absolutely resilient to $2(f +
1) - 1$ or fewer failures. A \phatraidcf{$C$}{$f$} group is likely resilient
to $2(f+1)$ failures and unlikely resistant to $2(f+1)+1$ failures.

The general form for the probability of \phatraidcf{$C$}{$f$} group failure
given $2(f + 1)$ node failures is:

\begin{align*}
  \Pr\left[\textrm{group failure} \,|\, \phatraidcf{C}{f}, n_{failure}=2(f+1)\right]
    =
  \frac{\binom{C}{2}\binom{2f+1}{f+1}^2}
       {\binom{C \cdot (2f+1)}{2(f+1)}}
\end{align*}

For \phatraidcf{6}{1} we calculate:

\begin{align*}
  P\left[\textrm{group failure} \,|\, \phatraidcf{6}{1}, n_{failure}=4\right]
    = \frac{135}{3000} \approx 0.05
\end{align*}

@@

\section{Future Work}

Extra parity bits!
